---
title: "p8105_hw6_lmp2234"
author: "Lisa Pardee"
date: "2024-11-30"
output: github_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)


```

```{r}
library(tidyverse)
library(modelr)
library(mgcv)
library(SemiPar)
library(p8105.datasets)
library(purrr)
set.seed(1)
```


# Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Bootstrapping 
```{r}
weather_boot_results =
 weather_df %>% 
  bootstrap(5000, id = "strap_number") %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy)
  ) %>%
  select(strap_number, results) %>%
  unnest(results) 
  
  weather_boot_results %>% 
  group_by(term) %>% 
  summarize(
    mean_est = mean(estimate), 
    sd_est = sd(estimate)
  )

```

Extracting R^2 from the bootstrap model 


```{r}
weather_boot_results = 
  weather_df %>% 
  bootstrap(5000, id = "strap_number") %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin, data = .x)),
    r2 = map_dbl(models, ~broom::glance(.x)$r.squared)
  )
  

```

Calculating log(Bo x B1)

```{r}
weather_boot_results = weather_boot_results %>% 
  mutate(
    coefs = map(models, broom::tidy),
    log_beta_product = map_dbl(coefs, ~{
      coefs = .x
      beta0 = coefs$estimate[coefs$term == "(Intercept)"]
      beta1 = coefs$estimate[coefs$term == "tmin"]
      log(beta0 * beta1)
    })
  )
```

Mapping the distribution of each estimate 

```{r}
library(ggplot2)

ggplot(weather_boot_results, aes(x = r2))+
  geom_density()+
  labs(title = "Bootstrap Distribution of R^2", x = "R2", y = "Frequency")
  
ggplot(weather_boot_results, aes (x = log_beta_product))+
  geom_density() +
  labs(title = "Boostrap Distribution of Log Beta Products", x = "Log Beta", y = "Frequency") 

```

Calculating Confidence Intervals for Both 
```{r}
r2_ci = weather_boot_results %>%
  summarize(
    ci_lower = quantile(r2, 0.025),
    ci_upper = quantile(r2, 0.975)
  ) %>%
  mutate(term = "R^2")

r2_ci

coef_ci = weather_boot_results %>% 
  unnest(coefs) %>% 
  group_by(term) %>% 
  summarize(
    ci_lower = quantile(estimate, 0.025),
    ci_upper = quantile(estimate, 0.975)
  )

coef_ci 
  
boot_confidence_intervals = bind_rows(coef_ci, r2_ci)

boot_confidence_intervals
```

# Problem 2 

```{r, message = FALSE, warning = FALSE}
library(readr)

homocide_df = 
  read_csv("homicide_data.csv") %>% 
  mutate(
    city_state = paste(city, state, sep = ","), 
    solved = if_else(disposition == "Closed by arrest", 1, 0), 
    victim_age = as.numeric(victim_age)) %>% 
  filter(
    !city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"),
    victim_race %in% c("White", "Black"), 
    !is.na(victim_age)
  )
```

```{r}
baltimore_fit = glm(solved ~ victim_age + victim_sex + victim_race, data = homocide_df, family = binomial())

coef(baltimore_fit)

baltimore_tidy = 
baltimore_fit |> 
  broom::tidy(conf.int = TRUE)  |> 
  select(term, estimate, conf.low, conf.high, p.value)

baltimore_tidy

male_female_odds = 
  baltimore_tidy %>% 
  filter(term=="victim_sexMale")

male_female_odds

```
Running glms for each city 

```{r}
library(purrr)

city_regressions = 
  homocide_df %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(
    glm_fit = map(data, ~glm(solved ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
    tidy_results = map(glm_fit, ~broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) %>% 
  unnest(tidy_results) %>% 
  filter(term == "victim_sexMale") %>%
  select(city_state, estimate, conf.low, conf.high, p.value)

city_regressions
```
```{r}
city_regressions %>% 
  arrange(estimate) %>% 
  mutate(
    city_state = factor(city_state, levels = city_state)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  labs( title = "Esimated ORs and CIs for Solving Homocides by City", x = "City", y = "Male to Female Victims Odds Ratio") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```
This plot displays that cities such as Albuquerque, Freston, and Stockton have higher odds ratios of male to female victims for solving homicides.

#Problem 3 

```{r}
birthweight_df = 
  read_csv("birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex) ,
    frace = as.factor(frace) ,
    malform = as.factor(malform) ,
    mrace = as.factor(mrace)) %>% 
      drop_na()
```

Looking at the data 
Fitting a model that includes factors such as mother's weigth at delivery (pounds),  

```{r}
birthweight_model <-
  lm(bwt ~ delwt + gaweeks + ppbmi + ppwt + delwt*gaweeks , data = birthweight_df)

summary(birthweight_model)

```
Residuals and Predictions
```{r}
birthweight_df <- birthweight_df %>% 
  add_predictions(birthweight_model, var = "pred_bwt") %>% 
  add_residuals(birthweight_model, var = "residuals")

library(ggplot2)
ggplot(birthweight_df, aes(x=pred_bwt, y = residuals)) +
  geom_point() + 
  geom_hline(yintercept = 0, color = "red")+
  labs(title = "Residuals vs Fitted Values", x = "Predicted Birthweight", y = "Residuals") 
  
```


Fitting the regression for a model of length at birth and gestational age as predictors 

```{r}
birthweight_df |> 
  lm(bwt ~ blength + gaweeks, data = _) |> 
  broom::tidy()  |> 
  knitr::kable()
```

Fitting a regression model for head circumference, length, sex, and interactions 
```{r}
birthweight_df |> 
  lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex + bhead*blength + bhead*babysex + blength*babysex, data = _) |> 
  broom::tidy()  |> 
  knitr::kable(digits = 4)
```

Cross-Validation

```{r}
set.seed(1)

cv_df <- crossv_mc(birthweight_df, n = 100)|> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Fitting Models & Extracting RMSEs

```{r}
cv_df <- cv_df %>% 
  mutate(
    model1 = map(train, ~ lm(bwt ~ delwt + gaweeks + ppbmi + ppwt + delwt * gaweeks, data = .)),
    model2 = map(train, ~ lm(bwt ~ blength + gaweeks, data = .)),
    model3 = map(train, ~ lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex + bhead*blength + bhead*babysex + blength*babysex, data = .))
  ) |> 
    mutate(
    rmse_model1 = map2_dbl(.x = model1, .y = test, ~ rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(.x = model2, .y = test, ~ rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(.x = model3, .y = test, ~ rmse(model = .x, data = .y))
  )
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(), 
    names_to = "model", 
    values_to = "rmse", 
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(title = "RMSE Distribution for Each Model", x = "Model", y = "RMSE")
```
Compared to the other models, my model has the worst prediction accuracy while model 3 has the most accurate prediction accuracy since the RMSE is lower. 

